---
title: 'Reproducible Research Assignment #1'
author: "Pey-Ran Che"
date: "Saturday, July 18, 2015"
output: html_document
---


## Loading and preprocessing the data
### Dataset: Activity monitoring data

1. Load the data (i.e. read.csv())
```{r}
x <- read.csv(file="activity.csv", header=TRUE, sep=",")
head(x)
```



2. Process/transform the data (if necessary) into a format suitable for your analysis

   Not Necessary



## What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.

1. Calculate the total number of steps taken per day
```{r}
y <- aggregate(list(steps=x$steps), by=list(date=x$date), FUN=sum, na.rm = TRUE)
head(y)
```


2. If you do not understand the difference between a histogram and a barplot, research 
   the difference between them. Make a histogram of the total number of steps taken each day
```{r, echo=TRUE}
hist(y$steps, breaks=10)
```


3. Calculate and report the mean and median of the total number of steps taken per day
```{r, echo=TRUE}
mean(y$steps)
median(y$steps)
```



## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number 
   of steps taken, averaged across all days (y-axis)
```{r}
y2 <- aggregate(list(steps=x$steps), by=list(time=x$interval), FUN=mean, na.rm = TRUE)
head(y2)
```

```{r, echo=TRUE}
plot(y2$time, y2$steps, type="l", xlab="", pch=".", ylab="average number of steps taken")
```


2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number 
   of steps?
```{r, echo=TRUE}
y2[y2$steps==max(y2$steps),]
```


## Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). 
The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number 
   of rows with NAs)
```{r}
sum(is.na(x$steps))
```


2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not 
   need to be sophisticated. For example, you could use the mean/median for that day, or the mean for 
   that 5-minute interval, etc.

   Missing value (NA) will be replaced by the mean (rounded) for that 5-minute interval.
   

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
t <- x
for(i in 1:nrow(t)) { 
        if (is.na(t[i,][1])){
                s <- t[i,][3]
                u <- y2[y2$time==s$interval,]
                t[i,][1] <- round(u$steps)
        }
}
head(t)
```


4. Make a histogram of the total number of steps taken each day and Calculate and report the 
   mean and median total number of steps taken per day. Do these values differ from the estimates 
   from the first part of the assignment? What is the impact of imputing missing data on the estimates 
   of the total daily number of steps?
   
```{r}
r <- aggregate(list(steps=t$steps), by=list(date=t$date), FUN=sum, na.rm = TRUE)
head(r)
```

```{r, echo=TRUE}
hist(r$steps, breaks=10)
```
Compare with previous histogram, there is not much difference except the intervals with small step counts have decreased while the middle ones have increaded.


```{r, echo=TRUE}
mean(r$steps)
median(r$steps)
```

Both mean and median value in here greater than those values in step 1.


## Are there differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.


1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating 
   whether a given date is a weekday or weekend day.
   
   (a) Convert date to day of week
```{r}
o <- t
o$day1 <- weekdays(as.Date(o$date))
head(o)
```


   (b) Convert day of week to factor variable weekday/weekend
```{r}
weekdays0 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
o$day2 <- factor((o$day1 %in% weekdays0), 
                   levels=c(FALSE, TRUE), labels=c('weekend', 'weekday') )
head(o)
```


2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) 
   and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
   See the README file in the GitHub repository to see an example of what this plot should look like 
   using simulated data.

```{r}
y3 <- aggregate(list(steps=o$steps), by=list(time=o$interval, day=o$day2), FUN=mean, na.rm = TRUE)
head(y3)
```

```{r, echo=TRUE}
library(ggplot2)
ggplot(data = o, aes(x = interval, y = steps))  +
        geom_line(color = "blue") +
        facet_grid(day2 ~ .)
```

